---
title: "homework3_qqnorm2"
author: "Denglinsui"
date: "2019/10/4"
output: html_document
---
## Analysis of the flight delay data

### Pre-treat data
We library relevant packages, read the data and transform them into data.table form firstly.
```{r dealdata, message=FALSE}
library(MASS)
library(data.table)

flight <- read.csv("/jddata1/homework/flights.csv",colClasses=c("YEAR" = "character", "DEPARTURE_TIME" = "character",
                                              "MONTH" = "character", "WHEELS_OFF" = "character",
                                              "DAY" = "character", "ORIGIN_AIRPORT" = "character",
                                              "SCHEDULED_DEPARTURE" = "character", "WHEELS_ON" = "character",
                                              "SCHEDULED_ARRIVAL" = "character", "ARRIVAL_TIME" = "character"))
flight <- as.data.table(flight)

#Date manipulate(departure_time)
# flight$departure_time <- ISOdatetime(flight$YEAR, flight$MONTH, 
#                                      flight$DAY,as.integer(substr(flight$SCHEDULED_DEPARTURE,1,2)),
#                                      as.integer(substr(flight$SCHEDULED_DEPARTURE,3,4)),0)
flights2 <- flight
```

Unlike before, we select all analyzable columns without preferences and 16 columns are selected.
```{r dealdata2}
#Select col
select_col <- c("MONTH", "DAY", "DAY_OF_WEEK", 
                "AIRLINE", "ORIGIN_AIRPORT", "DESTINATION_AIRPORT", "SCHEDULED_DEPARTURE", 
                "DEPARTURE_TIME", "DEPARTURE_DELAY", "TAXI_OUT", "SCHEDULED_TIME", 
                "ELAPSED_TIME", "AIR_TIME", "DISTANCE", "TAXI_IN", "ARRIVAL_DELAY")
flights2 <- flights2[, select_col, with=F]

# nan
flights2 <- na.omit(flights2, invert = F)

flights2 <- flights2[is.na(as.numeric(flights2$ORIGIN_AIRPORT)),] #numeric airport id. should be deleted.

flights2_dim <- dim(flights2)
print(flights2_dim)
```

We want to convert arilines, airports, months, days of month, days of week and etcs into dummy variables. However, there are nearly 300 airports and if we take them into considerationm, the memory is full, so we delete it. (However, we can try to classify the airport to solve this problem.)

The departure time influence the delay largely, whereas the departure time doesn't have meanings itself. Therefore, we split the departure time into three classes: 0~8, 8~16, 16~24.

```{r dummyvariable}
# dummy variable
library(nnet)
month_mat <- class.ind(flights2$MONTH)
colnames(month_mat) <- paste("month_", colnames(month_mat), sep = '')

day_mat <- class.ind(flights2$DAY)
colnames(day_mat) <- paste("day_", colnames(day_mat), sep = '')

day_of_week_mat <- class.ind(flights2$DAY_OF_WEEK)
colnames(day_of_week_mat) <- paste("day_of_week_", colnames(day_of_week_mat), sep = '')

airline_mat <- class.ind(flights2$AIRLINE)
colnames(airline_mat) <- paste("airline_", colnames(airline_mat), sep = '')

# 300 airport, memory is full!!!
# ori_airport_mat <- class.ind(flights2$ORIGIN_AIRPORT)
# colnames(ori_airport_mat) <- paste("origin_airport_", colnames(ori_airport_mat), sep = '')
# 
# des_airport_mat <- class.ind(flights2$DESTINATION_AIRPORT)
# colnames(des_airport_mat) <- paste("des_airport_", colnames(des_airport_mat), sep = '')

## SCHEDULED_DEPARTURE
flights2$SCHEDULED_DEPARTURE <- as.numeric(flights2$SCHEDULED_DEPARTURE)
flights2$is_first_sch_dep <- 0
flights2[between(SCHEDULED_DEPARTURE, 0, 799), is_first_sch_dep := 1 ]

flights2$is_second_sch_dep <- 0
flights2[between(SCHEDULED_DEPARTURE, 800, 1599), is_second_sch_dep := 1]

flights2$is_third_sch_dep <- 0
flights2[between(SCHEDULED_DEPARTURE, 1600, 2359), is_third_sch_dep := 1]

## DEPARTURE_TIME
flights2$DEPARTURE_TIME <- as.numeric(flights2$DEPARTURE_TIME)
flights2$is_first_dep <- 0
flights2[between(DEPARTURE_TIME, 0, 799), is_first_dep := 1]

flights2$is_second_dep <- 0
flights2[between(DEPARTURE_TIME, 800, 1599), is_second_dep := 1]

flights2$is_third_dep <- 0
flights2[between(DEPARTURE_TIME, 1600, 2359), is_third_dep := 1]


flights2 <- cbind(flights2, month_mat, day_mat, day_of_week_mat,
                  airline_mat) 
# flights2 <- cbind(flights2, month_mat, day_mat, day_of_week_mat,
#                   airline_mat, ori_airport_mat, des_airport_mat) 

```

Finally, the tail of delay time is heavy, we use logarithm transformation to reduce the heavy tail. To retain the information as more as possible, we substruct the minimal of delay time and take the logarithm transformation.

```{r step1_step2}
#step1&2
flights0930_2 <- flights2[,-c(1:8,34:64)]
# flights0930_2 <- flights0930_2[sample(1:nrow(flights0930_2), 10000, replace = F), ]

# 测试的数据，100000个, 已经删去没用的变量
# write.csv(flights0930_2, "test_data")
# flights0930_2 <- read.csv("test_data.csv")

# step1, step2
min_departure <- min(flights0930_2$DEPARTURE_DELAY)
min_arrival <- min(flights0930_2$ARRIVAL_DELAY)
flights0930_2=flights0930_2[DEPARTURE_DELAY>0&ARRIVAL_DELAY>0]
```

```{r arrival}
qqnorm(flights0930_2$ARRIVAL_DELAY)
qqnorm(log(flights0930_2$ARRIVAL_DELAY-min_arrival))
```

```{r departure}
qqnorm(flights0930_2$DEPARTURE_DELAY)
qqnorm(log(flights0930_2$DEPARTURE_DELAY-min_departure))
```